A Novel Solution to Resonant Scattering with Self-Consistent Boundary Conditions

### 2022-04-12 ms.tex
Removing all instances of redundant significant figures in plots, captions, and text

Several other works have followed \citet{1973MNRAS.162...43H}. \citet{1990ApJ...350..216N} extends the solution to media of intermediate optical depth, including the effects of scattering in the Doppler core of the line. \citet{2006ApJ...649...14D} generalize the same problem to spherical geometry, as is used here. \textbf{\cite{2020MNRAS.497.3925L} generalize both the slab and sphere solutions to arbitrary power-law density and emissivity profiles.}  Each of these works, however, use the same surface boundary condition and ansatz as \citet{1973MNRAS.162...43H}.

\textbf{The motivation for including time-dependence in the transfer equation is to characterize the distribution of photon escape times, which is needed to calculate the energy deposited into the gas by an escaping photon. Additionally, steady-state solutions to this problem are not always sufficient to describe all the physics of \lya transport. Time-variable, optically-thick environments necessitate a time-dependent solution to include the dynamic effects of \lya transfer. These include the optical afterglow of gamma-ray bursts \citep{2010ApJ...716..604R} and \lya sources redshifted by cosmological expansion %, which may reveal information about their neutral hydrogen halos and the intergalactic medium 
\citep{2011MNRAS.418..853X}.}

@@

 \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{xinit_threepanel.pdf}
    \caption{The same as Figure \ref{fig:sol_mc_residual_0}, but for $\tau_0=10^7$ and $\rm x_s = 0.0$ (top panel), $6.0$ (middle panel), and $12.0$ (bottom panel). The optical depth at each of these source frequencies is $\tau_s = 10^7, 77,$ and $19$, respectively.} 
    \label{fig:sol_mc_xinit}
\end{figure}

 \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{xinit_threepanel_tau1e6.pdf}
    \caption{The same as Figure \ref{fig:sol_mc_xinit}, but at a lower optical depth $\tau_0 = 10^6$. The shift $x_s$ is a much larger fraction of the distance to the spectral peak $(a\tau_0)^{1/3}$, and thus the asymmetry in the spectrum is much larger. The optical depth at the source frequency is $\tau_s = 10^6$, $7.7$, and $1.9$ for $x_s=0.0, 6.0,$ and $12.0$, respectively. } 
    \label{fig:sol_mc_xinit_lowtau}
\end{figure}

@@

Figure \ref{fig:sol_mc_xinit} shows calculations performed for $x_s = 0.0, 6.0$, and $12.0$ and $\tau_0=10^7$. The asymmetry of the spectrum is slight for $\rm x_s = 6.0$ where $\tau(x_s)=77$, but is larger for $\rm x_s=12.0$ outside the line core where $\tau(x_s)=19$. It is seen here that the difference between the Monte Carlo data and $H_0$ becomes larger as $\rm x_s$ increases. Thus, for large $|x_s|$, inclusion of $H_{\rm bc}$ is more important.

@@

The eigenfunction's oscillatory forms have varying amplitudes which sum in Equation (\ref{eq:Jrsigmat}) to create the final form of the mean intensity. The largest contribution at late times always comes from the ($n=1, m=1$) lowest-order eigenfunction. Figure \ref{fig:jsoln} shows a set of eigenfunctions $J_{nm}(\sigma)$ to illustrate their relative scales for different $m$ at a fixed spatial eigenmode $n$. The overall scale of the $J_{nm}(\sigma)$ are set by the factor $E/(kR^3)$ with $E$ arbitrarily set to 1. For H atoms with $T=10^4$ K and $\tau_0=10^7$, an eigenfunction has typical size ${\sim} E a / \left(R^2 \Delta \right) = 10^{-37}$ in units of specific mean intensity times time. Additional terms add smaller-magnitude, faster-oscillating components that lead to higher accuracy upon summation with the lower-order terms in Equation (\ref{eq:Jrsigmat}). The oscillations of various modes must cancel in the Fourier sum, so many modes $m$ and $n$ are required for convergence to the solution. 
\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{gamma_nm.pdf}
    \caption{Decay rate vs. mode number. Overlapping solid lines are plotted for each $n$. The thickness of the lines decrease with $n$ in order to clearly show where they are in agreement. These numerically-obtained resonant frequencies are compared with the analytic expression in Equation (\ref{eq:gamma_nm}), shown as a dashed line. The parameters for the calculation are $\tau_0=10^7$ and $x_s=0.0$. The value of $n$ is given by the color bar on the right-hand side.}
    \label{fig:gamma_nm}
\end{figure}
The values of the $\gamma_{nm}$ can be described approximately with Equation (\ref{eq:gamma_nm}). Their values depend on $m$, $n$, and other physical parameters according to 
\be \label{eq:gamma_nm}
\gamma_{nm} = 2^{-1/3} \pi^{13/6} n^{4/3}\left(m-\frac{7}{8}\right)^{2/3}\frac{c}{R}(a\tau_0)^{-1/3}
\ee
as shown \textbf{by setting the denominator of Equation \ref{eq:a3/a2} to zero in WKB approximation of} Appendix \ref{app:wkb}. The power law in $m$ is weak, requiring up to $m=1000$ to reduce the scale of $\gamma_{nm}^{\ \ -1}$ by two orders of magnitude. When sweeping through to find resonances, Equation (\ref{eq:gamma_nm}) is used to set the scale of the sweep points $\gamma_j$ to ensure no $\gamma_{nm}$ are missed. The continuity and close agreement with the analytic expression shown in Figure \ref{fig:gamma_nm} indicate that the numerical solutions are accurate.

@@

\begin{figure}
    \centering
    \includegraphics{waittime.pdf}
    \caption{Wait-time distribution of escaping photons $P(t)$ vs. t  for $\tau_0=10^6$ and $10^7$, including the fitting function from Equation (\ref{eq:fitting_function}). A sum over 20 spatial eigenmodes and 500 frequency eigenmodes is labeled ``Eigenfunctions''. All calculations were performed with a monochromatic source of photons at line center ($x_s = 0.0$).}
    \label{fig:escape_time}
\end{figure}

@@

There are several methods that are commonly used to accelerate Monte Carlo radiation transfer calculations, including core skipping methods \citep{1968ApJ...153..783A,2002ApJ...567..922A} and hybrid diffusion methods \citep{2018MNRAS.479.2065S}, but these methods are not sufficient for our application since they do not \textbf{directly measure or sample} the amount of time photons spend in the line core, which is needed to tabulate the intensity. Another approach with wide application is modified random walk methods, such as those discussed in \citet{1984JCoPh..54..508F, 2009A&A...497..155M, 2010A&A...520A..70R}. In this approach, an outgoing photon is randomly sampled on the surface of the outgoing sphere by drawing its properties from distributions in outgoing frequencies, directions, and escape times, based on solutions to the diffusion equation. A method similar to this has been applied by \citet{2006ApJ...645..792T} to Lyman $\alpha$ transfer using the \citet{1990ApJ...350..216N} solution, but this solution of course does not utilize the frequency-dependent boundary condition at the surface of the sphere. Furthermore, to perform a full radiation hydrodynamic simulation with Monte Carlo acceleration, it will be necessary to calculate radiation forces due to \lya transfer. Similar calculations have been done in \citet{1976ApJ...208..286W} in plane-parallel geometry. However, these solutions are limited to optical depths below $2.5 \times 10^3$. For this work, it would be necessary to model line center optical depths of up to 1 million or more.

>>>>

Several other works have followed \citet{1973MNRAS.162...43H}. \citet{1990ApJ...350..216N} extends the solution to media of intermediate optical depth, including the effects of scattering in the Doppler core of the line. \citet{2006ApJ...649...14D} generalize the same problem to spherical geometry, as is used here. \textbf{\cite{2020MNRAS.497.3925L} generalize both the slab and sphere solutions to arbitrary power-law density and emissivity profiles. Each of these works, and several others \citep{2020ApJS..250....9S, 2021MNRAS.504...89T}, use either the same surface boundary condition and ansatz as \citet{1973MNRAS.162...43H}, or use a solution that does not handle the frequency-dependence of the boundary condition. Our novel steady-state solution involves a frequency-dependent correction to the boundary condition that fixes an excess at the spectral peaks as compared with Monte Carlo, which is present in many of the works cited above.}

\textbf{The motivation for including time-dependence in the transfer equation is to characterize the distribution of photon escape times, which is needed to calculate the energy deposited into the gas by an escaping photon. Additionally, steady-state solutions to this problem are not always sufficient to describe all the physics of \lya transport. Time-variable, optically-thick environments necessitate a time-dependent solution to include the dynamic effects of \lya transfer. These include the optical afterglow of gamma-ray bursts \citep{2010ApJ...716..604R} and \lya sources redshifted by cosmological expansion %, which may reveal information about their neutral hydrogen halos and the intergalactic medium 
\citep{2011MNRAS.418..853X}, among others.}

@@

 \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{xinit_threepanel.pdf}
    \caption{The same as Figure \ref{fig:sol_mc_residual_0}, but for $\tau_0=10^7$ and $\rm x_s = 0$ (top panel), $6$ (middle panel), and $12$ (bottom panel). The optical depth at each of these source frequencies is $\tau_s = 10^7, 77,$ and $19$, respectively.} 
    \label{fig:sol_mc_xinit}
\end{figure}

 \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{xinit_threepanel_tau1e6.pdf}
    \caption{The same as Figure \ref{fig:sol_mc_xinit}, but at a lower optical depth $\tau_0 = 10^6$. The shift $x_s$ is a much larger fraction of the distance to the spectral peak $(a\tau_0)^{1/3}$, and thus the asymmetry in the spectrum is much larger. The optical depth at the source frequency is $\tau_s = 10^6$, $7.7$, and $1.9$ for $x_s=0, 6,$ and $12$, respectively. } 
    \label{fig:sol_mc_xinit_lowtau}
\end{figure}

@@

Figure \ref{fig:sol_mc_xinit} shows calculations performed for $x_s = 0, 6$, and $12$ and $\tau_0=10^7$. The asymmetry of the spectrum is slight for $\rm x_s = 6$ where $\tau(x_s)=77$, but is larger for $\rm x_s=12$ outside the line core where $\tau(x_s)=19$. It is seen here that the difference between the Monte Carlo data and $H_0$ becomes larger as $\rm x_s$ increases. Thus, for large $|x_s|$, inclusion of $H_{\rm bc}$ is more important.

@@

The eigenfunction's oscillatory forms have varying amplitudes which sum in Equation (\ref{eq:Jrsigmat}) to create the final form of the mean intensity. The largest contribution at late times always comes from the ($n=1, m=1$) lowest-order eigenfunction. Figure \ref{fig:jsoln} shows a set of eigenfunctions $J_{nm}(\sigma)$ to illustrate their relative scales for different $m$ at a fixed spatial eigenmode $n$. The overall scale of the $J_{nm}(\sigma)$ are set by the factor $E/(kR^3)$ with $E$ arbitrarily set to 1. For H atoms with $T=10^4$ K and $\tau_0=10^7$, an eigenfunction has typical size ${\sim} E a / \left(R^2 \Delta \right) = 10^{-37}$ in units of specific mean intensity times time. Additional terms add smaller-magnitude, faster-oscillating components that lead to higher accuracy upon summation with the lower-order terms in Equation (\ref{eq:Jrsigmat}). The oscillations of various modes must cancel in the Fourier sum, so many modes $m$ and $n$ are required for convergence to the solution. 
\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{gamma_nm.pdf}
    \caption{Decay rate vs. mode number. Overlapping solid lines are plotted for each $n$. The thickness of the lines decrease with $n$ in order to clearly show where they are in agreement. These numerically-obtained resonant frequencies are compared with the analytic expression in Equation (\ref{eq:gamma_nm}), shown as a dashed line. The parameters for the calculation are $\tau_0=10^7$ and $x_s=0$. The value of $n$ is given by the color bar on the right-hand side.}
    \label{fig:gamma_nm}
\end{figure}
The values of the $\gamma_{nm}$ can be described approximately with Equation (\ref{eq:gamma_nm}). Their values depend on $m$, $n$, and other physical parameters according to 
\be \label{eq:gamma_nm}
\gamma_{nm} = 2^{-1/3} \pi^{13/6} n^{4/3}\left(m-\frac{7}{8}\right)^{2/3}\frac{c}{R}(a\tau_0)^{-1/3}
\ee
as shown \textbf{by setting the denominator of Equation \ref{eq:a3/a2} to zero in WKB approximation of} Appendix \ref{app:wkb}. The power law in $m$ is weak, requiring up to $m=1000$ to reduce the scale of $\gamma_{nm}^{\ \ -1}$ by two orders of magnitude. When sweeping through to find resonances, Equation (\ref{eq:gamma_nm}) is used to set the scale of the sweep points $\gamma_j$ to ensure no $\gamma_{nm}$ are missed. The continuity and close agreement with the analytic expression shown in Figure \ref{fig:gamma_nm} indicate that the numerical solutions are accurate.

@@

\begin{figure}
    \centering
    \includegraphics{waittime.pdf}
    \caption{Wait-time distribution of escaping photons $P(t)$ vs. t  for $\tau_0=10^6$ and $10^7$, including the fitting function from Equation (\ref{eq:fitting_function}). A sum over 20 spatial eigenmodes and 500 frequency eigenmodes is labeled ``Eigenfunctions''. All calculations were performed with a monochromatic source of photons at line center ($x_s = 0$).}
    \label{fig:escape_time}
\end{figure}

@@

There are several methods that are commonly used to accelerate Monte Carlo radiation transfer calculations, including core skipping methods \citep{1968ApJ...153..783A,2002ApJ...567..922A} and hybrid diffusion methods \citep{2018MNRAS.479.2065S}. Another approach with wide application is modified random walk methods, such as those discussed in \citet{1984JCoPh..54..508F, 2009A&A...497..155M, 2010A&A...520A..70R}. In this approach, an outgoing photon is randomly sampled on the surface of the outgoing sphere by drawing its properties from distributions in outgoing frequencies, directions, and escape times, based on solutions to the diffusion equation. A method similar to this has been applied by \citet{2006ApJ...645..792T} to Lyman $\alpha$ transfer using the \citet{1990ApJ...350..216N} solution, but this solution of course does not utilize the frequency-dependent boundary condition at the surface of the sphere. Furthermore, to perform a full radiation hydrodynamic simulation with Monte Carlo acceleration, it will be necessary to calculate radiation forces due to \lya transfer. Similar calculations have been done in \citet{1976ApJ...208..286W} in plane-parallel geometry. However, these solutions are limited to optical depths below $2.5 \times 10^3$. For this work, it would be necessary to model line center optical depths of up to 1 million or more.
